---
title: "Project_1_mussel"
author: "Me"
date: "24/10/2020"
output: html_document
---

```{r echo=FALSE, results='hide', message=FALSE}
# required libraries
library(dplyr)
library(tidyr)
library(dlookr)
library(outliers)
library(ggplot2)
library(ggpubr)
library(knitr)
library(markdown)
options(digits=5)
```
## How old is mussel?

In this work presented statistical analysis of size and age of the mussels. 

### Work data
Save work directory to 'My_way'.

```{r}
My_way <- "/home/stas/git/BI_Stat_2020/project_1_mussels//data/"
```

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = My_way)
```


Collect all data to one table using import Data function.
```{r echo=FALSE}
importData <- function(My_way){
  list_of_csv <- list.files(My_way, pattern = '*.csv') 
  merged_data <- lapply(list_of_csv, read.csv, header = TRUE)
  combined_data <- do.call(rbind, merged_data)
}
```

```{r }
df <- importData(My_way)
```

### EDA

1) Rename 'Sex..1...male..2...female..3...uvenil' 
2) Change type 'Sex..1...male..2...female..3...uvenil', 'Rings' и 'Length' 

```{r echo=FALSE}
str(df)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
df <- df %>%
  rename(
    Sex = Sex..1...male..2...female..3...uvenil.
  )
df$Sex <- factor(df$Sex, levels = c(1, 2, 3), labels = c('Male', 'Female', 'Uvenil'))

df$Rings <- as.numeric(df$Rings)
df$Length <- as.numeric(df$Length)

str(df)
```

```{r echo=FALSE}
na_tab <- colSums(is.na(df))
```

Our data contains `r sum(na_tab)` of NA. 

I deleted all samples with NA because there are not many of them. It is the most simple way, but not the best. 

```{r echo=FALSE}
Mussel_data <- na.omit(df)
```
Check outliers for each sexes.


```{r echo=FALSE}

Height_in_Sex <- ggboxplot(Mussel_data, x = 'Sex' , y = 'Height',
                           color = 'Sex', title = 'Height', ylab = FALSE, xlab = FALSE) + 
                              theme(legend.position = "none") +
                              rotate_x_text(angle = 45) +
                                theme(plot.title = element_text(hjust = 0.5))

Rings_in_Sex <- ggboxplot(Mussel_data, x = 'Sex', y = 'Rings',
                          color = 'Sex', title = 'Rings', ylab = FALSE, xlab = FALSE) + 
                            theme(legend.position = "none") +
                            rotate_x_text(angle = 45) + 
                            theme(plot.title = element_text(hjust = 0.5))

Length_in_Sex <- ggboxplot(Mussel_data, x = 'Sex', y = 'Length',
                           color = 'Sex', title = 'Length', ylab = FALSE, xlab = FALSE) + 
                            theme(legend.position = "none") +
                            rotate_x_text(angle = 45) +
                            theme(plot.title = element_text(hjust = 0.5))

Diameter_in_Sex <- ggboxplot(Mussel_data, x = 'Sex', y = 'Diameter',
                             color = 'Sex', title = "Diameter", ylab = FALSE, xlab = FALSE) + 
                              theme(legend.position = "none") +
                              rotate_x_text(angle = 45) +
                               theme(plot.title = element_text(hjust = 0.5))

Whole_weight_in_Sex <- ggboxplot(Mussel_data, x = 'Sex', y = 'Whole_weight',
                                 color = 'Sex', title = 'Whole weight', ylab = FALSE, xlab = FALSE) + 
                                    theme(legend.position = "none") +
                                     rotate_x_text(angle = 45) +
                                      theme(plot.title = element_text(hjust = 0.5))

Shucked_weight_in_Sex <- ggboxplot(Mussel_data, x = 'Sex', y = 'Shucked_weight',
                                   color = 'Sex', title = 'Shucked weight', ylab = FALSE, xlab = FALSE) + 
                                    theme(legend.position = "none") +
                                    rotate_x_text(angle = 45) +
                                    theme(plot.title = element_text(hjust = 0.5))

Viscera_weight_in_Sex <- ggboxplot(Mussel_data, x = 'Sex', y = 'Viscera_weight',
                                   color = 'Sex', title = 'Viscera weight', ylab = FALSE, xlab = FALSE) + 
                                      theme(legend.position = "none") +
                                      rotate_x_text(angle = 45) +
                                      theme(plot.title = element_text(hjust = 0.5))

Shell_weight_in_Sex <- ggboxplot(Mussel_data, x = 'Sex', y = 'Shell_weight',
                                 color = 'Sex', title = 'Shell weight', ylab = FALSE, xlab = FALSE) + 
                                  theme(legend.position = "none") +
                                  rotate_x_text(angle = 45) +
                                  theme(plot.title = element_text(hjust = 0.5))


Half_boxplotes <-  ggarrange(Height_in_Sex, Rings_in_Sex, Length_in_Sex, Diameter_in_Sex)

Weight_boxplotes <- ggarrange(Whole_weight_in_Sex, Shucked_weight_in_Sex, Viscera_weight_in_Sex, Shell_weight_in_Sex)

Half_boxplotes 
Weight_boxplotes
```

'Height' has two outliers also confirmed by Grubbs's test. They will be deleted from analysis.

```{r echo=FALSE}
grubbs.test(Mussel_data$Height)
Mussel_data <-  subset(Mussel_data, Height != max(Mussel_data$Height))
grubbs.test(Mussel_data$Height)
Mussel_data <-  subset(Mussel_data, Height != max(Mussel_data$Height))
```

```{r echo=FALSE}
Height_in_Sex <- ggboxplot(Mussel_data, x = 'Sex' , y = 'Height',
                           color = 'Sex', title = 'Height', ylab = FALSE) + 
                              theme(legend.position = "none") +
                              rotate_x_text(angle = 45) +
                                theme(plot.title = element_text(hjust = 0.5))
Height_in_Sex
```

Check interconnection between variables.

```{r echo=FALSE}
plot_correlate(Mussel_data)
```

### EDA results

1) Strong correlation between length and diameter 
2) Strong correlation between weight and diameter 
3) Juveniles are smaller than adults


### Main

Mean and sd for mussels from different sex
```{r echo=FALSE, warning=FALSE, message=FALSE}
mean_sd_for_Sex <- Mussel_data %>% 
  group_by(Sex) %>%
  summarise(Mean = mean(Length), Standart_deviation = sd(Length))

```
```{r echo=FALSE}
kable(mean_sd_for_Sex)
```

Calculate percent of mussels with Height less than 0.165

```{r echo=FALSE}
percent_0.165 <-  round(sum(Mussel_data$Height <= 0.165) / sum(Mussel_data$Height >= 0), 4) 
```

75.82% of mussels have Height less than 0.165


```{r echo=FALSE}
number_of_92 <- round((sum(Mussel_data$Length > 0) * 0.92), 0)
count_92_percentage <- Mussel_data %>% arrange(Length)
Eithn_data <- count_92_percentage[-c(1:number_of_92),]
```

Length value which greater than length of 92% of other mussels: `r min(Eithn_data$Length)`


### Create Length_z_scores

```{r echo=FALSE}
Mussel_data <- Mussel_data %>% mutate(Lenght_z_scores = scale(Length))
kable(Mussel_data[c(1:5),])
```

### Statistic tests

Compare mussels with 5 and 15 rings
```{r echo=FALSE}
diam_5 <- Mussel_data$Diameter[Mussel_data$Rings == 5]
diam_15 <- Mussel_data$Diameter[Mussel_data$Rings == 15]
testo <- t.test(diam_15, diam_5, alternative = 'greater')
testo
```

Mussels with 15 rings have greater diameter than mussels with 5 rings (t = `r testo$statistic`, df = 196, p-value < 2e-16)


### Diametr and Whole_weight

```{r echo=FALSE}
cor_test_Diam_Weight <- cor.test(Mussel_data$Diameter, Mussel_data$Whole_weight)
cor_test_Diam_Weight
```

Diameter and Whole_weight are correlated  (t = `r cor_test_Diam_Weight$statistic`, df = `r cor_test_Diam_Weight$parameter` , p-palue < 2e-16)